This thesis would not have been possible without the help from many,
especially my advisor, Prof.\ Jason Wright. His tireless mentoring and
encouragement provided strong support for the later years of my
graduate study. His intellectual brilliance, academic integrity, and
genuine altruism as an advisor and mentor has made him my role model
for the academic path.

I thank my families for their support in many many ways, especially my
husband, Dr.\ Enshi Xu, for his generous support and whine-free
devotion to the housework. I thank my daughter for truly understanding
the importance of my work to me (especially for her willingness to get
out of the house quickly in the mornings even when she did not want to but
had to go to daycare). 

I thank all my committee members, Profs.\ Larry Ramsey, Jim Kasting,
Suvrath Mahadevan, and Eric Ford for guidance and helpful
suggestions. Their brain storming ideas on job hunting had put me in
contact with the PFS team at the Carnegie Institute for Science and
eventually landed me my dream job as a Carnegie Fellow at DTM after
graduation.

I thank all the Penn State Astronomy faculty, staff, and students for
providing a warm home for me for the past eight years. I truly think
we are the best department on Earth. In particular, I would like to
thank Profs.\ Peter M\'esz\'aros and Niel Brandt for taking me in on
their projects, providing me financial support, and teaching me many
valuable skills as an astronomer. I would like to thank Profs.\ Robin
Ciadullo, Steinn Sigurdsson, and Don Schneider for their mentoring and
moral support. I thank my fellow graduate students, especially Megan
Comins, Prakash Arumugasamy, and Jason Young, for moral support.

Many collaborators have contributed to the work in this thesis. I am
in debt of Dr.\ John A.\ Johnson for providing his copy of the
California Planet Survey (CPS) Doppler code and for his generous
advice and help on Doppler reduction in general. I thank Dr.\ Debra
Fischer for her help with Doppler reduction in general, and for
inviting me to host the breakout session on telluric contamination at
the EPRV II workshop in Yale in July 2015. I thank Dr.\ Chad Bender
for his generous help with TERRASPEC and on modeling the telluric
lines. I am grateful for Dr.\ Bill Cochran and the UT Austin team for
providing archival HD 37605 data and for their collaboration on the
discovery of HD 37605$c$. I am very grateful for CPS for the same
reason, and also for providing many other data, such as Keck spectra
on HD 185144, which are very crucial for my work. I thank the HET
staff for helping us on the engineering run for characterizing modal
noise in HRS, and I am extremely thankful for the staff at McDonald
Observatory, especially David Doss and Coyne Gibson, for their help
with the TS12 runs. Many thanks to Drs.\ Gillian Nave and Stephen
Redman for performing the NIST FTS scan for the HET/HRS iodine
cell. Last but not least, I would like to thank Dr.\ Ming Zhao for his
work on TS12, as well as his generous help and advice throughout our
overlapping years at Penn State and even beyond.

I am very grateful for the funding provided by NSF (grant AST-1211441),
NASA NESSF, CEHW, and NAI at Penn State for the work in this thesis.

