

\begin{small}

\begin{center}
{\Large Jason E. Young}\\
525 Davey Lab\\
University Park, PA 16802\\
(814) 863-7948\\
\link{jyoung@astro.psu.edu}\\
\link{www.astro.psu.edu/~jyoung}\\


\end{center}
\bigskip


\topic{Research Expertise}\\
\begin{itemize}
 \item low surface brightness galaxies
 \item quasars / AGN
 \item star-forming and active galaxies
 \item optical, UV, near-, mid-, and far-IR photometry
 \item UV, optical, and near-IR spectroscopy
 \item computational/numerical analysis
 \item C, C++, Python, Java, IRAF CL, Bash
\end{itemize}

\topic{Education}

\quadrupleitem{Pennsylvania State University}{University Park, PA}
{Doctor of Philosophy in Astronomy and Astrophysics}{December 2012}

\quadrupleitem{University of Arizona}{Tucson, AZ}
{Bachelor of Science in Astronomy and Physics, minor in Latin}{May 2005}


\topic{References}\\
\begin{itemize}
\item Chris Palma,\qquad\,\, cpalma@psu.edu,\qquad\quad\, (814) 865-2255, 525 Davey Lab
\item Caryl Gronwall,\quad\, caryl@astro.psu.edu,\qquad (814) 865-2918, 525 Davey Lab
\item Michael Eracleous, mce@astro.psu.edu,\qquad\, (814) 863-6041, 525 Davey Lab
\item Jane Charlton,\quad\quad charlton@astro.psu.edu,\, (814) 863-6040, 525 Davey Lab
\item Donald Schneider,\,\, dps7@psu.edu,\qquad\qquad\, (814) 863-9554, 525 Davey Lab
\end{itemize}

\pagebreak
\topic{Research Experience}

\doubleitem{Pennsylvania State University}{University Park, PA}

\tripleitem{Researcher}{June 2013 - present}
{Principle Investigator on a program to determine the star-formation histories of low surface brightness galaxies.}

\tripleitem{Research Assistant/Associate}{August 2012 - June 2013}
{Search for close binary supermassive black holes using HST STIS and COS spectra of SDSS quasars.}

\tripleitem{Research Assistant}{May 2009 - August 2012}
{Detection and analysis of quasar host galaxies in WFPC2 and NICMOS images using narrow- and medium-band filters. Developed a novel technique to perform PSF subtraction on continuum and emission-line images of eight PG quasars. Using the emission-line residual images, star-formation rates and morphologies were estimated for each quasar host galaxy.}

\tripleitem{Research Assistant}{June 2005 - August 2012}
{Analysis of multi-wavelength properties of star-forming galaxies in the Kitt Peak International Spectroscopic Survey (KISS). Photometry of star-forming galaxies in the UV, optical, and near-, mid-, and far-IR was compared with star-formation rates in search of novel star formation rate indicators and bulk properties of star-forming galaxies.}

\tripleitem{Research Assistant}{May 2007 - May 2009}
{Detection and photometry of star-forming regions and dwarf galaxies in tidal debris and interacting galaxies in Compact Galaxy Groups using ACS images. Stellar evolution models were used to determine age and mass for each object.}

\doubleitem{University of Arizona}{Tucson, AZ}

\tripleitem{Research Assistant}{September 2002 - May 2005}
{Helped discover twelve high redshift (z$\approx$6) quasars by following up photometric candidates from the Sloan Digital Sky Survey using near-IR photometry and spectroscopy.}

\topic{Posters \& Presentations}
\begin{itemize}
\item Young, J. E., Eracleous, M., Shemmer, O., Gronwall, C., Netzer, H., Lutz, D., Sturm, E., \& Ciardullo, R. ``In Search of Star-Forming Regions in the Host Galaxies of Nearby Quasars'' 2010, 215th AAS Meeting
\item Young, J. E., Gronwall, C., Charlton, J. C., Durrell, P. R., Elmegreen, D., Gallagher, S. C., Johnson, K., \& Chandar, R., ``ACS Photometry of Tidal Debris in HCG 31'' 2008, 211th AAS Meeting
\item Young, J. E., Gronwall, C., Rosenberg, J. L., \& Salzer, J. J., ``Spitzer IRAC Observations of Nearby Star-Forming Galaxies and Implications for Star-Formation Rates'' 2007, 209th AAS Meeting
\end{itemize}

\pagebreak
\topic{Talks}
\begin{itemize}
\item  Young, J. E. ``Quasar Host Galaxies: What's Really Going On?'', lunch talk, Space Telescope Science Institute, January 2013
\item  Young, J. E. ``Star Formation Across Galactic Environments'', dissertation talk, Winter AAS Meeting, January 2013
\item  Young, J. E. ``Quasar Host Galaxies: What's Really Going On?'', lunch talk, Goddard Space Flight Center, December 2012
\item  Young, J. E. ``Quasar Host Galaxies: What's Really Going On?'', lunch talk, Cornell, October 2012
\item  Young, J. E. "In Search of Star Formation in Quasar Host Galaxies," Penn State departmental talk, Spring 2011
\item  Young, J. E., Eracleous, M., Gronwall, C., Shemmer, O., Netzer, H., Sturm, E., \& Ciardullo, R. 2011, in Bulletin of the America
 Astronomical Society, Vol. 43, American Astronomical Society Meeting Abstracts 217, 421.04-+
\item  Young, J. E. "Quasar Host Galaxies: The Story of a Hidden Train Wreck," departmental talk, Fall 2009
\item  Young, J. E. "C++: Object Oriented Programming," Graduate Student Association presentation, Summer 2008
\end{itemize}

\topic{Teaching Experience}

\doubleitem{Pennsylvania State University}{University Park, PA}

\tripleitem{Lecturer}{August 2013 - present}
{Taught an introductory astronomy course focusing on the sky and planets as well as an astronomy overview course.}


\tripleitem{Instructor}{August 2005 - August 2012}
{Taught four sections of an elementary astronomy laboratory for undergraduate non-majors.}

\tripleitem{Rooftop Telescope Instructor/Operator}{August 2005 - Present}
{Taught a rooftop telescope activity for undergraduate non-majors.}

\tripleitem{Teaching  Assistant}{Fall 2010, Fall 2011}
{Assisted with the teaching of an advanced observational techniques course for undergraduate astronomy majors.}

\doubleitem{University of Arizona}{Tucson, AZ}

\tripleitem{Telescope Operator}{August 2003 - May 2005}
{Operated the Steward Observatory telescope for introductory astronomy lab activities.}
\\\pagebreak
\doubleitem{Julia-Keen Elementary}{Tucson, AZ}

\tripleitem{Tutor}{August 2002 - May 2004}
{Tutored local elementary school students in math and reading during in-class activities as well as after school.}

\topic{Professional Societies}

\doubleitemnobold{American Astronomical Society}{2005 - Present}
\doubleitemnobold{Graduate Women in Science (National Nomination Committee)}{2010 - Present}
\doubleitemnobold{Penn State Graduate Student Association (Vice President, Committee Chair)}{2008 - 2011}

\topic{Outreach Experience}

\doubleitem{Flandrau Science Center}{Tucson, AZ}

\tripleitem{Planetarium Console Operator}{2000 - 2005}
{Gave sky tours and operated planetarium.}

\tripleitem{Planetarium Maintenance Assistant}{2003 - 2005}
{Performed technical maintenance of planetarium projection equipment.}

\tripleitem{Science Education Assistant}{2002 - 2005}
{Performed science demonstrations and experiments for visiting groups.}

\tripleitem{Telescope Operator}{2002 - 2005}
{Operated the 16-inch observatory for the general public.}


\doubleitem{Pennsylvania State University}{University Park, PA}

\tripleitem{Volunteer}{2005 - Present}
{\begin{itemize}
 \item AstroFest, $\sim$2,000 visitors/year, Volunteer \& Organizer, 2005-2012
 \item Space Day / Exploration Day, $\sim$1,000 visitors/year, Volunteer \& Organizer, 2005-2012
 \item Astronomy Open House, $\sim$200 visitors/year, Volunteer \& Organizer, 2005-Present
 \item Shaver's Creek Environmental Center, Teaching Volunteer \& Special Guest Lecturer, 2008-Present
 \end{itemize}
}

\tripleitem{Night Sky Radio Program}{2008 - 2011}
{Weekly guest on a public radio program; provided night sky updates and viewing suggestions.}

\end{small}
